---
title: "Assignment2"
author: "Samuel Andrews and Mikolaj Wieczorek"
date: "2/3/2020"
output: word_document
---
```{r include=FALSE}
library(car)
```

# Problem 1

```{r}
Concrete = read.csv("Concrete.csv")
set.seed(1)
sam = sample(1:1030, size = floor(.6666*1030), replace = F)
Concrete.trans = Concrete
str(Concrete.trans)
summary(Concrete.trans)
```


a)	Fit a MLR model with all variables in their original scales using the training data.  Summarize this model and discuss any model deficiencies. 

```{r fig.height=15, fig.width=15}
# BlastFurn, FlyAsh, and Superplast have zeros
lm1 = lm(Strength~., data = Concrete.trans[sam,])
par(mfrow=c(2,2))
plot(lm1)
```

There is a little bit of constant variance - some transformation could help it.
High leverage pulling the model to the right. We also have some outliers - but they are not that terrible in this case.
Normality does not seem to be affected. There also seems to be some curvature present in the model.


```{r}
vif(lm1)
VIF(lm1)
```

Multicollinearity does not seem to be an issue in this model as the variance inflation factor is not larger than 10 for predictors.

```{r}
#Check the training and validation sets nrow()
nrow(Concrete.trans[sam,])
nrow(Concrete.trans[-sam,])
```

b)	Use the model from part (a) to predict the response value using the validation data and compute the prediction accuracy (RMSEP,MAEP,MAPEP) of these predictions by comparing the actual compression strengths of the concrete samples in the validation set.  

```{r}
y = Concrete.trans$Strength[-sam]
ypred = predict(lm1, newdata = Concrete.trans[-sam,])
results = PredAcc(y, ypred)
```

RMSEP = 
MAEP = 
MAPEP = 





d)	Fit a MARS model to the training data with degree = 1 (i.e. no interactions).  Use the internal cross-validation features of the earth() function to choose the “best” MARS model with degree = 1.  Again predict the compression strength of the concrete samples in the validation set in the original scale and compute RMSEP, MAEP, and MAPEP.  How does this compare to the models in part (a) and (c)?  (10 pts.)


We will use our variable transformed dataset.

```{r}
#install.packages("earth")
require(earth)

Concrete.mars = earth(Strength~.,data = Concrete.trans[sam,], degree = 1 )
plot(Concrete.mars)
plotmo(Concrete.mars)
```
```{r}
summary(Concrete.mars)

```
Upon a quick runthrought of the Mars Method, we selected 16 terms from 8 predictors and got a GRSq of 0.854, which is impressive compared to the time required to due a more traditional Multiple Linear regression in order to achieve similar results. Still, we have yet to check the model using internal Cross-validation methods or after adjusting paramters, so we aren't done yet.
```{r}
Concrete.mars = earth(Strength~.,data = Concrete.trans[sam,], degree = 1, keepxy = T, nfold = 10, ncross = 30 )

plot(Concrete.mars)

```

Based on 10 Kfolds, crossing 30 times, we seem to picking from 16 to 19 terms for our model when degree is 1, which is similar to our intial. We also should look at how variables are contributing to the model, just to make sure nothing seems out of wack.
```{r}
evimp(Concrete.mars, trim = FALSE)
```
As indicated by the summary, all but one of the variables are used. This is intereting  given what we learned through our MLR efforts, as CourseAgg tended to be a poor predictor simply in general and was decied to be left out for those models as well. Otherwise, like we found earlier age and cement seem to be big ticket contributors, while superplasta and flyash are weaker as indivdals. Although we should note that water is much more effective here, likely due to its structure being more akin to mappign vai checkmar functions than linear models with polynomial terms.

For the sake of being thorough, we will also attempt this process again after goading the model to utilize less variables to see if it has any effect on the GRSq. Due to it already using almost everything, it makes little sense to force more, but a simpler model may potentially prove more effective. However, given what we know about out variables thus far, it seems unlikely that much success will be found as truncating from seven likely decrease our model.



```{r}
Concrete.mars2 = earth(Strength~.,data = Concrete.trans[sam,], degree = 1, nprune = 5, nfold = 10, keepxy = T, ncross = 30 )
plot(Concrete.mars2)

```

```{r}
plotmo(Concrete.mars2)
```


```{r}
summary(Concrete.mars2)
```


```{r}
evimp(Concrete.mars2, trim = FALSE)
```

As expected, our GRSq lowered by a little over .10,  as well as other similar measures. While our k-fold cross validation does more concretly setttle on a specific nunmber of terms (5), the massive loss in predictive ability likely isn't worth it. 

Overall, while minor gains have been made in our favored metrics, it was decided that these small improvments were not worth taking hits to the conistancy, therefore the inital model was chosen despite some of its less desireable features.

```{r}
#0.6

ypredtransformed = predict(Concrete.mars, newdata = Concrete.trans[-sam,])

ypred = ypredtransformed^(1/0.6)

PredAcc(Concrete[-sam,]$Strength, ypred)


```

RMSEP
================
37.61542 

MAE
================
32.3783 

MAPE
================
92.37868 



Next, we will try with degree 2, allowing interactions to be considered.

```{r}
Concrete.mars.deg2 = earth(Strength~.,data = Concrete.trans[sam,], degree = 2, keepxy = T, nfold = 10, ncross = 30 )

plot(Concrete.mars.deg2)

```
Across 30 run-throughs, it seems to sway between 17 and 19 terms.

```{r}
plotmo(Concrete.mars.deg2)
```
```{r}
summary(Concrete.mars.deg2)
```
```{r}
evimp(Concrete.mars.deg2, trim = FALSE)
```


With a GRSq of just of 0.841, with interaction, we get a slightly worse model. We also are using less variables, despite there being more terms used. CourseAgg being removed makes sense, as it was not very useful in MLR and had a structure that was difficult to work with. 

Overall, adding interaction did not make a huge difference in our model's performance, however we may be able to widen the gap a bit if we adjust some parameters.

```{r}
Concrete.mars2.deg2 = earth(Strength~.,data = Concrete.trans[sam,], degree = 2, keepxy = T, nfold = 10, ncross = 30 , nk = 30, nprune = 10)

plot(Concrete.mars2.deg2)
```





```{r}
plotmo(Concrete.mars2.deg2)
```



```{r}
summary(Concrete.mars2.deg2)
```


```{r}
evimp(Concrete.mars2.deg2, trim = FALSE)
```
Pruning at 10 and leaving a max terms of 30 gives us an ultimatley weaker model, although we have manged to remain with measures above 0.8 whilst only using 10 terms. However, we the amount of varition across  folds is astounding with anywhere from 10 to 29 terms being used from run-through to run-through. Due to such inconsistancy, this iteration will likely be shelved.

Next we'll try using no prunes but forcing a higher minimum, though this may risk overfitting the model. 

```{r}
Concrete.mars3.deg2 = earth(Strength~.,data = Concrete.trans[sam,], degree = 2, keepxy = T, nfold = 10, ncross = 30 , nk = 25)

summary(Concrete.mars3.deg2)

```
```{r}

evimp(Concrete.mars3.deg2, trim = FALSE)
```
```{r}
plot(Concrete.mars3.deg2)
```
When forcing a higher minimum term count, we get a model with the best GRSq yet at about 0.861 that hovers between 15 and 25. Once again, it suffers from inconsistany, which is a considerable negative. for the next iteration, we will try this same model but re-add in nprune to seee if less greed can increase it further.

```{r}
Concrete.mars4.deg2 = earth(Strength~.,data = Concrete.trans[sam,], degree = 2, keepxy = T, nfold = 10, ncross = 30 , nk = 25, nprune = 21)

summary(Concrete.mars4.deg2)

```


```{r}
evimp(Concrete.mars4.deg2, trim = FALSE)
```


```{r}
plot(Concrete.mars4.deg2)
```

We get pretty much the same result other than minor variation not really worth discussing. As a result, this model continues to suffer from all of the same pitfalls as before with no real benefit. The only thing left to try that may help is reducing nk outright, to see if forcing a simpler model may fit something better after pruning.
```{r}
Concrete.mars5.deg2 = earth(Strength~.,data = Concrete.trans[sam,], degree = 2, keepxy = T, nfold = 10, ncross = 30 , nk = 15, nprune = 13)

summary(Concrete.mars5.deg2)
```

```{r}
evimp(Concrete.mars5.deg2, trim = FALSE)

```


```{r}
plot(Concrete.mars5.deg2)
```
While this version is much more consistant, it fails to be as effective as just our initial degree 2 model, with it having a GRSq about 3% lower. For one final iteration, we will try to make a previous model more conistant with pruning by focusing on less prunes.

```{r}
Concrete.mars4.deg2.md = earth(Strength~.,data = Concrete.trans[sam,], degree = 2, keepxy = T, nfold = 10, ncross = 30 , nk = 25, nprune = 15)

summary(Concrete.mars4.deg2.md)

```
```{r}
evimp(Concrete.mars4.deg2.md, trim = FALSE)
```

```{r}
plot(Concrete.mars4.deg2.md)
```
Alas, we once again arrive at the conondrum of minor improvements in the GRSq (about +0.2) in exchange for less consistency derrived from the internal CV methods. Favoring conistancy over vartion, we wil once again go with the defualt output from the mars function, as it seems like the best candiate among the ones we have manged to produce in contrast to it.

```{r}
#Predict
ypredtransformed2 = predict(Concrete.mars.deg2, newdata = Concrete.trans[-sam,])

ypred2 = ypredtransformed2^(1/0.6)

PredAcc(Concrete[-sam,]$Strength, ypred2)
```
RMSEP
36.99874 

MAE
31.79073 

MAPE
91.20575 



